resistanceModelNotesAndy

### 2015-03-12 resistanceModelNotesAndy.Rmd
progress notes

@@

@@

1. copy R functions out of `malaria_code_andy` into their own files in an R folder
1. add roxygen documentation for functions
1. later, structure as R package, maybe move to a new github repo

>>>>

```
#working version of Beths code
#run it outside of this project because it changes wd
source("C:\\Dropbox\\Ian and Andy\\andy\\malaria\\Beth code\\malaria_code_beth.r")
```

@@

**12/3/15** 9.30-14 4.5hrs 3-5 6-7 7.5hrs

The model generates 3 matrices : results, genotype, fitness
1. Results : freq of R allele at each loci in each sex and linkage disequilibrium of R allele in each sex, per generation
2. Genotype : frequencies of each of the ten genotypes, per generation
3. Fitness : fitness scores of each genotype/niche combination (table 4. of Main Document)

When a file is input with multiple scenarios, each of the three matrices is stored in a list, where scenario number gives position of the matrix in the list.

results.list : 
fitness.list : 
genotype.list :

Can I put these into a single list so that they can be returned from a single function ?

done ~ put input object creation into a function
But there is still a fair bit of code that reads parameter values out of the input matrix and into named variables.

I might be able to restructure these collections of single variables into arrays.
e.g. for the exposure levels of m&f

```{r}
#to create an array a[sex][locus1][locus2]
sex <- c("F","M")
locus1 <- c("0","A","a")
locus2 <- c("0","B","b")
dimnames1 <- list( sex=sex, locus1=locus1, locus2=locus2 )
dim1 <- sapply(dimnames1, function(x) length(x))
a <- array(0,dim=dim1, dimnames=dimnames1)
a
length(a)
#to access one element
a['M','a','b']
#to check that male exposures total 1
sum(a['M',,])
```

```
  ## Exposure levels of males and females to each insecticide niche ##
  # males
  a.m_00 <- input[8,i]

  a.m_a0 <- input[9,i]
  a.m_A0 <- input[10,i]

  a.m_0b <- input[11,i]
  a.m_0B <- input[12,i]

  a.m_ab <- input[13,i]
  a.m_AB <- input[14,i]

  a.m_Ab <- input[15,i]
  a.m_aB <- input[16,i]

  #a.m <- sum(a.m_00, a.m_a0, a.m_A0, a.m_0b, a.m_0B, a.m_ab, a.m_AB, a.m_Ab, a.m_aB)
  #if ( a.m != 1 ){		 
  #	print( paste("Error in male exposures: must total one: ", a.m) )

  #	}else{
  #		print( paste( "Male exposures total 1: ", a.m ))
  #		}

  # females
  a.f_00 <- input[17,i]

  a.f_a0 <- input[18,i]
  a.f_A0 <- input[19,i]

  a.f_0b <- input[20,i]
  a.f_0B <- input[21,i]

  a.f_ab <- input[22,i]
  a.f_AB <- input[23,i]

  a.f_Ab <- input[24,i]
  a.f_aB <- input[25,i]
```

I should check that what we are doing fits in with what others are doing. We may be able to capitalise on existing tools.

[popgen on CRAN](http://cran.r-project.org/web/packages/popgen/popgen.pdf)
seems a bit old and doesn't do a huge amount, no vignette.

[gstudio](http://cran.r-project.org/web/packages/gstudio/) looks more promising. e.g. it defines locus objects.
[gstudio documentation](http://dyerlab.github.io/gstudio/)
A locus object is set like this, but I'm not sure if this would help us.
It does have some stuff on drift at end of help.
```
require(gstudio)
loc <- locus(c("C", "A"))
loc
```

pegas provides functions for the analysis of allelic data and of haplotype data from DNA sequences.
It requires and complements two other R-packages: ape and adegenet.
[pegas](http://ape-package.ird.fr/pegas.html)
[pegas data structures](http://ape-package.ird.fr/pegas/DefinitionDataClassesPegas.pdf)
includes a class called loci. : An object of class "loci" is a data frame where rows represent individuals
and columns are loci and optional additional variables.
Again I can't quite see how it would be useful for us.

There is a [HardyWeinberg](http://cran.r-project.org/web/packages/HardyWeinberg/vignettes/HardyWeinberg.pdf) package on CRAN. "The HardyWeinberg package consists of a set of tools for analyzing diallelic
genetic markers, and is particularly focused on the graphical representation of their (dis)equilibrium condition in various ways."

There is a CRAN [genetics task view](http://cran.r-project.org/web/views/Genetics.html).

The [genetics package](http://cran.r-project.org/web/packages/genetics/genetics.pdf) which also has a locus class, seems to be more detailed, about chromosomes etc. and probably not of use to us.

#### Structure of the results matrices

```{r}
max_gen <- 2 #just as exampel

# Set up results matrix - prints overall freq of R and S allele per locus per sex, LD and overall allele freq (i.e. 1)
results <- matrix ( nrow = max_gen, ncol = 11 )
colnames( results ) <- c( "Gen", "m.R1", "m.R2", "m.LD","f.R1", "f.R2", "f.LD", "M", "F", "dprime", "r2" )

# set up fitness by niche matrix - records fitness scores for each niche for each genotype
fitness <- matrix ( nrow = 10, ncol = 9, c(rep(0,90)))
colnames(fitness) <- c( "-,-", "a,-", "A,-", "b,-", "B,-", "a,b", "A,B", "A,b", "a,B" )
rownames(fitness) <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
						"RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
						"RR1SS2", "RR1RS2", "RR1RR2")

# set up genotype matrix - records frequencies of each of the 9 two locus genotypes each generation
genotype <- matrix( nrow=max_gen, ncol=11 )
colnames(genotype) <- c("gen", "SS1SS2", "SS2RS2", "SS1RR2", 
						"RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
						"RR1SS2", "RR1RS2", "RR1RR2")

```

Somehow it seems I've broken it, I'm getting NAs in results matrix.

The input files are different (but same structure) between Beths version that works and my new version ...
calibration is 100 in my failing version and 1012 in Beths, i think this is because I'm reading in the csv when I shouldn't.

I seem not to get to line 889 in my version
  	# male
  	f.m.SS1SS2 <- genotype.freq[1,]

Then suddenly it seemed to start working when I put a browser in.
It's something tricky around line 711, if it doesn't get to 889 it doesn't fill in the results matrix ??

Aha! It seems to have been this without the comments before the tilde's that stoppped it from working
```
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 		
  ### Loop to run the model from the initial conditions generated above ####
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
```

@@

1. check what the tilde does in R
1. copy the 'run' bit out of `malaria_code_andy.r` into a function 
1. later, structure as R package, maybe move to a new github repo
1. re-look at R popgen packages when I've got more of a handle on what we are doing.
